---
title: "Homework 5" 
author: "Ainsley Lotito" 
date: "November 12, 2025" 
output: html_document 
--- 

# Looking at the `data` subdirectory 

*Write a piece of code that lists the files you saved in the `data` subdirectory.* 

```{r}
list.files("../data")
```

# Mapping a city  

*Pick one city in the data. Create a map showing the locations of the homicides in that city, using the `sf` framework discussed in class. Use *`tigris` *to download boundaries for some sub-city geography (e.g., tracts, block groups, county subdivisions) to show as a layer underneath the points showing homicides. Use different facets for solved versus unsolved homicides and different colors to show the three race groups with the highest number of homicides for that city (you may find the* `fct_lump` *function from forcats useful for this).*

## Importing data 

To start off, let's read in the data and take a look at the cities we may choose from. 

```{r}
library(tidyverse)
library(forcats)
library(tigris)

homicides <- read.csv("../data/homicide-data.csv") 
unique(homicides$city)
```

I think it would be fun to use Baton Rouge. Next, I'm going to create an object containing just the observations for Baton Rouge. 

```{r}
baton_rouge <- homicides %>% 
  filter(city == "Baton Rouge")  
dim(baton_rouge)
baton_rouge %>% 
  slice(1:3)    

#chaning to an sf object  
library(sf)
baton_rouge_sf <- st_as_sf(baton_rouge, coords = c("lon", "lat")) %>% 
  st_set_crs(4269)
baton_rouge_sf %>% slice(1:3)
dim(baton_rouge_sf)
```

Making a map of baton rouge. I need to pull some tracts, block groups, and county subdivisions from tigris. 
The FIPS code for East Baton Rouge Parish is 22033, and West is 22121. I can therefore select by the countyFP of 033 and 121.

```{r}
library(tigris)
la_tracts <- tracts(state = "LA", year = 2023) #tracts first
ba_tracts <- la_tracts %>% 
  filter(COUNTYFP %in% c("033", "121")) 

la_blocks <- block_groups(state = "LA", year = 2023)
ba_blocks<- la_blocks %>% 
  filter(COUNTYFP %in% c("033", "121")) 

ggplot() +
  geom_sf(data = ba_tracts, fill = NA, color = "black") + # Census Tracts
  geom_sf(data= ba_blocks, fill = NA, color = "gray") + # Census Blocks
  geom_sf(data = baton_rouge_sf, aes(color= disposition))

```